The Conjunctive Disjunctive Graph Node Kernel for Disease Gene Prioritization

### 2017-10-28 neurocomputing.tex
F

\begin{abstract}
Biological relations encoded in data sources are naturally represented in the form of graphs. As a consequence, a high number of graph-based biological learning systems have been proposed and shown promissing results. In graph-based learning systems, one of the key points which determines their performance is the denifition of graph node similarity measure. Node similarities are normally measured by graph node kernels. However, most available graph node kernels do not show high discriminative capacity since they share two common limitations. First they are based on the diffusion phenomenon which does not effectively exploit the nodes' configuration. Second they are not able to proceed the auxiliary information associated to graph nodes.

In this paper, we propose an efficient graph node kernel that not only effectively exploits nodes' context, but also process with additional information which available on the graph nodes. Empirical evaluation on disease gene prioritization show that our proposed graph node kernel shows the state of the art concerning graph node kernels.

@@

\begin{keyword}
Graph node kernels, graph decomposition, disease gene prioritization
\end{keyword}
\end{frontmatter}

\linenumbers
\section{Introduction}
The release of advanced technologies is one of the main reasons for the
revolution in various scientific research fields. In Biological and Medical
domain, modern technologies are making it not only easier but also more
economical than ever to undertake experiments and creating applications.
As a consequence, a vast amount of biological data in terms of volume
and type is generated through scientific experiments, published literature,
high-throughput experiment technology, and computational analysis. This
huge quantity of data are saved as biological datasets and made discoverable
through web browsers, application programming interfaces, scalable search
technology and extensive cross-referencing between databases. Biological
databases normally contain information about gene function, structure, lo-
calization, clinical effects of mutations and similarities of biological sequences and structures.

@@

@@

2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/1 & \textbf{88.8/1} & 85.0/3 & 84.1/4 \\

@@

\hline
$\overline{AUC}$ & 66.6	& 63.5 & 67.8 &	67.6 & 73.2 & \textbf{77.8} & 73.6 & 75.8 \\

$\overline{Rank}$ & 6.09 & 7.09 &  5.36 & 4.82 & 4.45 & \textbf{1.64} & 4.09 & 2.36 \\
 \hline 
\end{tabular}
\label{fig:cdnk-biogps}
\end{table}

@@

\hline
$\overline{AUC}$ & 62.6	& 70.1 & 75.2 & 75.3 & 76.5 & 77.8 & 77.4 &	\textbf{78.7}
 \\
$\overline{Rank}$ & 7.73 & 6.82 & 4.55 & 4.09 & 3.91 & \textbf{2.27} & 3.18 & 2.45 \\
 \hline 
\end{tabular}
\label{fig:cdnk-pathways}
\end{table}
\section{Conclusions}
We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results. Moreover, we have also introduced the way to integrate ``side'' information in form of real valued vectors when it is available on graph to get even better performance of CDNK. In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.
%\section{The Elsevier article class}
%
%\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
%
%\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
%
%\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
%\begin{itemize}
%\item document style
%\item baselineskip
%\item front matter
%\item keywords and MSC codes
%\item theorems, definitions and proofs
%\item lables of enumerations
%\item citation style and labeling.
%\end{itemize}
%
%\section{Front matter}
%
%The author names and affiliations could be formatted in two ways:
%\begin{enumerate}[(1)]
%\item Group the authors per affiliation.
%\item Use footnotes to indicate the affiliations.
%\end{enumerate}
%See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
%
%\section{Bibliography styles}
%
%There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
%
%Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
\section*{Funding}
This work was supported by the University of Padova, Strategic Project BIOINFOGEN.
\section*{References}

>>>>

\begin{abstract}
Due to advances in data acquisition technologies (such as next generation sequencing), large amounts of information on biological systems are being encoded in relational data bases at an increasing rate. State of the art systems capable to deal with such structured data at scale,  include graph approaches based on the notion of diffusion kernels. In these systems, information is propagated in a discounted way according to the topology of the data network. A disadvantage of these approaches is that the specific configuration of the neighborhood of each node is ignored. Moreover it would be beneficial to integrate, when available, side information, encoded as arbitrary feature vectors associated to nodes or edges.
%Biological relations encoded in data sources are naturally represented in the form of graphs. As a consequence, a high number of graph-based biological learning systems have been proposed and shown promising results. In graph-based learning systems, one of the key points which determines their performance is the definition of graph node similarity measure. Node similarities are normally measured by graph node kernels. However, most available graph node kernels do not show high discriminative capacity since they share two common limitations. First they are based on the diffusion phenomenon which does not effectively exploit the nodes' configuration. Second they are not able to proceed the auxiliary information associated to graph nodes.
In this paper, we propose a graph decomposition approach that can exploit the information present in the local topology and that can efficiently incorporate vectorial side information. We consider the challenging problem of disease gene prioritization and show that we can achieve  state-of-the-art results.

@@

\begin{keyword}
Graph node kernels, disease gene association
\end{keyword}
\end{frontmatter}

\linenumbers
\section{Introduction}

The release of advanced technologies is one of the main reasons for the
revolution in various scientific research fields. In Biological and Medical domain, modern technologies are making it not only easier but also more economical than ever to undertake experiments and creating applications. As a consequence, a vast amount of biological data in terms of volume and type is generated through scientific experiments, published literature, high-throughput experiment technology, and computational analysis. This huge quantity of data are saved as biological datasets and made discoverable through web browsers, application programming interfaces, scalable search technology and extensive cross-referencing between databases. Biological databases normally contain information about gene function, structure, localization, clinical effects of mutations and similarities of biological sequences and structures.

@@

\subsection{Parameter Impact}
In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 and on the BioGPS network. For each parameter, first fix the values of all other parameters, but allow the value of this papameter to change. More precisely, we use the optimal tuple $C = 5, d^* = 1, D = 20, P = 15$ for the fixed of parameters' values. We let the changes of $C$ in $\lbrace 4, 5, 6, 7, 8, 9\rbrace$, $d^*$ in $\lbrace 1, 2, 3, 4\rbrace$, $D$ in $\lbrace 8, 10,  15, 20, 30, 40, 50\rbrace$ and real vector size, $P$ in $\lbrace 8, 10,  15, 20, 30, 50, 70, 90\rbrace$

@@

2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/2 & \textbf{88.8/1} & 83.2/4 & 84.8/3 \\

@@

\hline
$\overline{AUC}$ & 66.6	& 63.5 & 67.8 &	67.6 & 73.2 & \textbf{77.8} & 73.5 & 75.9 \\

$\overline{Rank}$ & 6.09 & 7.09 &  5.36 & 4.82 & 4.45 & \textbf{1.64} & 4.18 & 2.27 \\
 \hline 
\end{tabular}
\label{fig:cdnk-biogps}
\end{table}

@@

\hline
$\overline{AUC}$ & 62.6	& 70.1 & 75.2 & 75.3 & 76.5 & 77.8 & 77.4 &	\textbf{78.7}
 \\
$\overline{Rank}$ & 7.73 & 6.82 & 4.55 & 4.09 & 3.91 & \textbf{2.27} & 3.18 & 2.45 \\
 \hline 
\end{tabular}
\label{fig:cdnk-pathways}
\end{table}

\newpage
\begin{figure}[!tbp]
  \centering
  \begin{minipage}[b]{0.47\textwidth}
    \includegraphics[width=\textwidth]{cliques.png}
    \caption{Performance with C changes}
    \label{fig:c-change}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.47\textwidth}
    \includegraphics[width=\textwidth]{distance.png}
    \caption{Performance with d* changes}
    \label{fig:d-change}
  \end{minipage}
\end{figure}

\begin{figure}[!tbp]
\centering
  \begin{minipage}[b]{0.47\textwidth}
    \includegraphics[width=\textwidth]{degree.png}
    \label{fig:D-change}
    \caption{Performance with D changes}
  \end{minipage}
  \begin{minipage}[b]{0.47\textwidth}
    \includegraphics[width=\textwidth]{vec_size.png}
    \label{fig:vs-change}
    \caption{Performance with P changes}
  \end{minipage}  
\end{figure}

\section{Conclusions}
We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results. Moreover, we have also introduced the way to integrate ``side'' information in form of real valued vectors when it is available on graph to get even better performance of CDNK. In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.
%\section{The Elsevier article class}
%
%\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
%
%\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
%
%\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
%\begin{itemize}
%\item document style
%\item baselineskip
%\item front matter
%\item keywords and MSC codes
%\item theorems, definitions and proofs
%\item lables of enumerations
%\item citation style and labeling.
%\end{itemize}
%
%\section{Front matter}
%
%The author names and affiliations could be formatted in two ways:
%\begin{enumerate}[(1)]
%\item Group the authors per affiliation.
%\item Use footnotes to indicate the affiliations.
%\end{enumerate}
%See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
%
%\section{Bibliography styles}
%
%There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
%
%Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
\section*{Funding}
This work was supported by the University of Padova, Strategic Project BIOINFOGEN.
\section*{References}
